### 2024-02-07 ms.qmd
Updates

In this sense, prenatal Glucocorticoids (GCs) - hormones related to organisms' response to stress [@sapolsky_how_2000] - and prenatal thermal environment are known to influence learning abilities in different taxa [see @lemaire_prenatal_2000; @zhu_prenatal_2004; @amiel_hotter_2012; @amiel_egg_2014; @crino_corticosterone_2014; @abayarathna_effects_2020]. For instance, some studies have demonstrated that prenatal stress and high prenatal GC levels impair learning [@lemaire_prenatal_2000; @zhu_prenatal_2004; @farrell_developmental_2015-learn], while others showed diverse effects depending on factors like subjects' sex or the nature of the learning task [@szuran_water_1994; @crino_corticosterone_2014-learn; @farrell_developmental_2015-learn; @bebus_associative_2016]. Similarly, some experiments have shown significant impacts of prenatal temperature on learning in ectotherms [@amiel_hotter_2012; @amiel_egg_2014; @dayananda_incubation_2017; @abayarathna_effects_2020]. For example, high incubation temperatures have been linked with faster learning rates in skinks [@amiel_hotter_2012; @amiel_egg_2014], while velvet geckos incubated at temperatures over their natural range learn slower than those incubated within the natural thermal limits [@abayarathna_effects_2020]. In this vein, the effect of prenatal temperature appears to be linked to alterations in neural structure and metabolic activity [@coomber_independent_1997; @sakata_neural_2000; @amiel_effects_2017; @beltran_are_2021] that share some similarities with those resulting from prenatal increased stress or GC levels [@lemaire_prenatal_2000; @zhu_prenatal_2004; @du_dynamic_2009]. This suggests that prenatal GCs and temperature can act on the same physiological mechanisms and, thus, both could interact to shape individual variation in learning abilities [@noble_developmental_2018]. Furthermore, GCs can play an pivotal role in determining vertebrate responses to elevated temperatures [@Crino_2023] potentially fostering natural interactions between temperature and GCs. However, despite the proximate similarities of prenatal GCs and temperature effects, and the potential role of GCs in vertebrates' response to temperature, our understanding of how these two factors interact remains incomplete.  

@@

Our results suggest that the early environment manipulation did not affect learning in both skinks; we found that the estimated learning slopes were not different between treatments or species. However, learning slopes were lower when the blue feeders were the correct choice compared to those assigned to the group 'Red' for _L. delicata_, but not for _L. guichenoti_. This difference could be due to a potential bias towards blue in the initial choice, as we found that the estimated probability of choosing the correct feeder in the first trial was higher than expected by chance when the correct feeder was blue (see Supplementary Material). [**EXAMPLES WITH BIASES TOWARDS BLUE**]
Nevertheless, when lizards assigned to blue or red were analysed separately, the between treatments comparissons remain the same in both species. Furthermore, the non-significant trends we observed were not consistent between species or colour assigned. As such, we conclude that, overall, lizards learned at the same rate regardless of the early environment manipulation.

- Effect of CORT: what do others study say? Effect of sex, effect of task (brain region), effect of dose and time of exposure (discuss).
- Effects of temperature: what do others study say? Effect of tasks and region

Finally, we did not see any significant difference in learning rates when both species were compared. This result is in line with previous studies that found no differences in learning abilities between _L. delicata_ and _L. guichenoti_ [@bezzina2014does]. However, in the experiment conducted by @bezzina2014does the authors did not find any evidence of learning in neither _L. delicata_ or _L. guichenoti_. In this regard, both learning tasks were conducted in a similar way and using the same cognitive domain, so the differences in the results could be due to the different experimental designs or the complexity of stimuli discrimination. Nevertheless, to test cognitive difference between these two species, further studies should be conducted using different cognitive domains. 
In conclusion,

>>>>

In this sense, prenatal Glucocorticoids (GCs) - hormones related to organisms' response to stress [@sapolsky_how_2000] - and prenatal thermal environment are known to influence learning abilities in different taxa [see @lemaire_prenatal_2000; @zhu_prenatal_2004; @amiel_hotter_2012; @amiel_egg_2014; @crino_corticosterone_2014; @abayarathna_effects_2020]. For instance, some studies have demonstrated that prenatal stress and high prenatal GC levels impair learning [@lemaire_prenatal_2000; @zhu_prenatal_2004; @farrell_developmental_2015-learn], while others showed diverse effects depending on factors like subjects' sex or the nature of the learning task [@szuran_water_1994; @crino_corticosterone_2014-learn; @farrell_developmental_2015-learn; @bebus_associative_2016]. Similarly, some experiments have shown significant impacts of prenatal temperature on learning in ectotherms [@amiel_hotter_2012; @amiel_egg_2014; @dayananda_incubation_2017; @abayarathna_effects_2020]. For example, high incubation temperatures have been linked with faster learning rates in skinks [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014], while velvet geckos incubated at temperatures over their natural range learn slower than those incubated within the natural thermal limits [@dayananda_incubation_2017; @abayarathna_effects_2020]. In this vein, the effect of prenatal temperature appears to be linked to alterations in neural structure and metabolic activity [@coomber_independent_1997; @sakata_neural_2000; @amiel_effects_2017; @beltran_are_2021] that share some similarities with those resulting from prenatal increased stress or GC levels [@lemaire_prenatal_2000; @zhu_prenatal_2004; @du_dynamic_2009]. This suggests that prenatal GCs and temperature can act on the same physiological mechanisms and, thus, both could interact to shape individual variation in learning abilities [@noble_developmental_2018]. Furthermore, GCs can play an pivotal role in determining vertebrate responses to elevated temperatures [@Crino_2023] potentially fostering natural interactions between temperature and GCs. However, despite the proximate similarities of prenatal GCs and temperature effects, and the potential role of GCs in vertebrates' response to temperature, our understanding of how these two factors interact remains incomplete.  

@@

Our results suggest that the early environment manipulation did not affect learning; we found that the estimated learning slopes were not different between treatments or species. However, learning slopes were lower when the blue feeders were the correct choice compared to those lizards assigned to the group 'Red' for _L. delicata_, but not for _L. guichenoti_. This difference could be due to a potential bias towards blue in the initial choice, as we found that the estimated probability of choosing the correct feeder in the first trial was higher than expected by chance when the correct feeder was blue (see Supplementary Material). Moreover, those treatments and species whose learning slopes do not differ significantly from zero (see Table 1), are associated to highest bias (see Supplementary Material). Because there was a possibility that the bias towards the blue feeder was something acquired during training, we compared the light spectrum of the ramps used in the associative task and the white ones used during habituation, but we did no see any relevant similarity between the white and blue ramps (see Supplementary Material). As such, we consider this bias to not be a derivate from our experimental design. The species confidence hypothesis posits that animals should bias their attention toward colours they are familiar with, and empirical studies show that, for instance, dark and light blue T-shirts were associated with lower flight initiation distances and higher capture rates in Western fence lizards where blue is used in intraspecific communication [@putman2017fear]. However, to the best of our knowledge, noone has reported a colour bias in _L. delicata_ or _L. guichenoti_ before; and it is not considered to be involved in intraspecific communication [@torr1996patterns, @chapple_biology_2014]. Still, we recommend to select carefully the colours employed in associative tasks with these species, especially _L. delicata_, for future cognitive tasks employing visual stimuli. Nevertheless, when lizards assigned to blue or red were analysed separately, the between treatments comparissons remain the same in both species. Furthermore, the non-significant trends we observed were not consistent between species or colour assigned. As such, we conclude that, overall, lizards learned at the same rate regardless of the early environment manipulation.  
As part of our predictions, we expected prenatal CORT to impair learning, as other studies have shown [@lemaire_prenatal_2000; @zhu_prenatal_2004; @farrell_developmental_2015-learn; @bebus_associative_2016]. However, in other experiments early-life increases in GCs or induced stress did not affect learning abilities or even improve them [@szuran_water_1994; @crino_corticosterone_2014-learn; @farrell_developmental_2015-learn; @bebus_associative_2016]. In this veins, GCs are known to show an inverted 'U-shaped' relationship with performance on some learning task, with low levels of GCs being beneficial for learning and memory, and high levels being detrimental [@du_dynamic_2009; @mcewen_brain_2012]. It can be arguable whether the absence of differences in learning slopes between treatments could be due to the fact that the levels of CORT we used were not high or sustained through time enough to impair learning, or that the dose did not reach the embryo. However, we believe this to be unlikely since we found that the levels of CORT we used affected other morphological and physiological traits (non-published data). In contrast, testing only one cognitive domain could have masked potential differences in learning abilities between treatments. For instance, European starlings males stressed during early stages of development performed worse in a visual associative learning task, but show no differences with control birds when the stimuli tested was acoustic [@farrell_developmental_2015-learn]; while Scrub-jays that had lower CORT levels as nestlings performed better on an associative learning test as adults, but not on a reversal learning task [@bebus_associative_2016]. GCs action involve genomic and nongenomic mechanisms that implicate different types of receptors that can be distributed unevenly in the various regions of the brain [@mcewen_brain_2012], and may be related to the learning performance in distinct tasks. As such, we consider that studying different cognitive domains could provide a more comprehensive understanding of the effects of early-life stress on learning abilities in these species.  
Regarding incubation temperature, we predicted hot-incubated lizards to perform better in our task then those incubated at 23ºC, since most of the studies show enhanced learning abilities with higher incubation temperatures [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014]. Furthermore, those studies that found the opposite pattern, were only when the hot treatment was above the natural thermal range of the species [@dayananda_incubation_2017; @abayarathna_effects_2020]. However, our results do not align with these previous experiments and our predictions. In contrast, we did not find any effect of temperature in _L. delicata_ or _L. guichenoti_. In general, the incubation temperature effects observed on learning performance are associated with higher neural density in certain areas [@amiel_effects_2017], but these effects can be region-dependent [@coomber_independent_1997; @sakata_neural_2000; @amiel_effects_2017]. Hence, as ocurred with CORT, further studies using different cognitive domains should be used to understand the effects of incubation temperature on learning abilities in these species.  
Finally, we did not see any significant difference in learning rates when both species were compared. This result is in line with previous studies that found no differences in learning abilities between _L. delicata_ and _L. guichenoti_ [@bezzina2014does]. However, in the experiment conducted by @bezzina2014does the authors did not find any evidence of learning in neither _L. delicata_ or _L. guichenoti_. In this regard, both learning tasks were conducted in a similar way and using the same cognitive domain, so the differences in the results could be due to the different experimental designs or the complexity of stimuli discrimination. Nevertheless, to test cognitive difference between these two species, further studies should be conducted using different cognitive domains.    
In conclusion, we found that neither prenatal CORT, temperature, or their interaction affected learning abilities when _L. delicata_ and _L. guichenoti_ were subjected to a colour-associative task. However, we believe studying different cognitive domains is necessary to understand the effects of early-life stress and incubation temperature on learning abilities in these species. In addition, we found a bias towards blue that could affect the results found in cognitive tests, and we therefore recommend to select carefully the colours employed when employing visual stimuli in these two species, particularly _L. delicata_.
